Faster generic IND-CCA secure KEM using encrypt-then-MAC

### 2024-10-11 etm-transform/manuscript/main.tex
Getting ready for CHES submission

\subsection{Our contribution}
In this paper, we propose a generic IND-CCA secure KEM using techniques adapted from DHIES. We call our proposed scheme the encrypt-then-MAC transformation due to the conceptual similarity to the namesake symmetric encryption technique for achieving authenticated encryption. Our scheme mainly differs from DHIES in its versatility and input requirement. Whereas the IND-CCA security of DHIES reduces specifically to the Gap Diffie-Hellman assumption, the chosen-ciphertext security of the encrypt-then-MAC KEM reduces more generally to the OW-PCA security \cite{DBLP:conf/ctrsa/OkamotoP01} of the input scheme (we will show in section \ref{sec:main-results} that the Gap Diffie-Hellman assumption implies a special case of OW-PCA security). In addition, we propose that because each call to encapsulation samples a fresh PKE plaintext, the encrypt-then-MAC KEM can be instantiated with one-time secure MAC such as Poly1305 (Abdalla et al originally proposed to use HMAC and CBC-MAC, which are many-time secure MAC but less efficient than one-time MAC, see section \ref{sec:choosing-a-message-authenticator}), which greatly improves the computational efficiency of the scheme.

@@

\textbf{Security.} In Section \ref{sec:main-results}, we prove that if the input PKE is one-way secure against chosen-plaintext attack (OW-PCA) and the MAC is one-time existentially unforgeable, then the output KEM is IND-CCA secure. The IND-CCA security of the output KEM reduces tightly to the OW-PCA security of the input PKE, and non-tightly to the unforgeability of the MAC, though the non-tightness will likely not be a weakness because the construction makes it easy to plug in MAC of any desired security level.

\textbf{Implementation and performance comparison.} In Section \ref{sec:implementation}, we instantiate an instance of encrypt-then-MAC KEM using the PKE subroutines defined in ML-KEM \cite{FIPS203}. For clarity's sake we call this instantiation $\mlkemplus$. Compared with the full KEM routine in ML-KEM, which is based on some variant of the modular Fujisaki-Okamoto KEM transformation, $\mlkemplus$ demonstrates enormous runtime efficiency gain in decapsulation with only minimal performance penalty in encapsulation runtime and ciphertext size. On an AMD EPYC 9R14 CPU, $\mlkemplus$ (combined with Poly1305) achieves between 72\%-80\% reduction of CPU cycles needed for decapsulation while only incurring 2\%-7\% increase of CPU cycles needed for encapsulation. See Table \ref{tbl:cpu-cycles-summary} for a summary of performance comparison of individual routines against ML-KEM.

@@

    \caption{$\text{ML-KEM}^+$ achieves large performance improvements in decapsulation at the cost of minimal penalty in encapsulation performance and ciphertext size}\label{tbl:cpu-cycles-summary}
\end{table}

Since the decapsulation efficiency comes at the cost of increasing encryption runtime and ciphertext size, we also measured the round trip time of key exchange protocols with various modes of authentication. When compared to ML-KEM, $\mlkemplus$ achieves 24\%-28\% reduction of round trip time in unauthenticated key exchange (KE), 29\%-35\% in unilaterally authenticated key exchange (UAKE), and 35\%-48\% reduction in mutually authenticated key exchange (AKE). See Table \ref{tbl:rtt-summary} for a summary of round trip times.

@@

The modular Fujisaki-Okamoto KEM transformation is remarkably successful because of the simplicity of its construction, the tightness of the security bound, and the proven (though non-tight) security against quantum adversaries. It was adopted by many submissions to NIST's post-quantum cryptography competition, including Kyber \cite{DBLP:conf/eurosp/BosDKLLSSSS18}, Saber \cite{DBLP:conf/africacrypt/DAnversKRV18}, FrodoKEM \cite{DBLP:conf/ccs/BosCDMNNRS16}, and classic McEliece \cite{classicmceliecespec} among others.

However, the Fujisaki-Okamoto transformation is not perfect. Among its shortfalls are two problems caused by its use of \emph{de-randomization} and \emph{re-encryption} for achieving chosen-ciphertext security. \begin{itemize}
    \item \textbf{Computational inefficiency}. The decapsulation routine needs to re-encrypt the decryption to ensure ciphertext has not been tempered with. For input PKE whose encryption routine carries substantial computational cost, such as most lattice-based cryptosystems, re-encryption slows down decapsulation significantly.
    \item \textbf{Side-channel vulnerability}. Re-encryption also introduces side-channels that can leak information about the decrypted PKE plaintext. As demonstrated in \cite{ueno2022curse}\cite{ravi2019generic}\cite{DBLP:journals/tches/TanakaUXITH23}, these side-channels can be converted into efficient plaintext-checking attacks that can fully recover the secret key in fewer than 1000 traces.
\end{itemize}

@@

\subsection{Public-key encryption scheme}
\textbf{Syntax.} A public-key encryption scheme $\pke(\keygen, \encrypt, \decrypt)$ is a collection of three routines defined over some plaintext space $\mathcal{M}$ and some ciphertext space $\mathcal{C}$. $(\pk, \sk) \leftsample \keygen()$ is a randomized routine that returns a keypair. The encryption routine $\encrypt: (\pk, m) \mapsto c$ encrypts the input plaintext under the input public key. The decryption routine $\decrypt: (\sk, c) \mapsto m$ decrypts the input ciphertext under the input secret key. Where the encryption routine is randomized, we denote the randomness by $r \in \mathcal{R}$, where $\mathcal{R}$ is called the coin space. The decryption routine is assumed to always be deterministic. Some decryption routines can detect malformed ciphertext and output the rejection symbol $\bot$ accordingly.

@@

Where the expectation is taken with respect to the probability distribution of all possible keypairs $(\pk, \sk) \leftsample \texttt{PKE.KeyGen()}$. For many lattice-based cryptosystems, including ML-KEM, decryption failures could leak information about the secret key, although the probability of a decryption failure is low enough that classical adversaries cannot exploit decryption failure more than they can defeat the underlying lattice problem. On the other hand, a quantum adversary may be able to exploit decryption failure in reasonable runtime by efficiently searching through all possible inputs using Grover's search algorithm. For that, ML-KEM made slight modifications in its KEM construction to prevent quantum adversary from precomputing large lookup table. We refer readers to \cite{avanzi2019crystals} and \cite{DBLP:conf/eurosp/BosDKLLSSSS18} for details of the mitigation techniques, though decryption failure is not a focus of this paper.

\textbf{Security.} The security of public-key encryption is conventionally discussed within the context of adversarial games played between a challenger and an adversary \cite{DBLP:conf/stoc/GoldwasserM82}. Each game is defined by the goal and the capabilities of the adversary.

@@

    \begin{minipage}[t]{0.4\textwidth}
        \begin{algorithm}[H]
            \caption*{\texttt{OW-ATK} Game}
            \begin{algorithmic}[1]
                \State $(\pk, \sk) \leftsample \keygen(1^\lambda)$
                \State $m^\ast \leftsample \mathcal{M}$
                \State $c^\ast \leftsample \encrypt(\pk, m^\ast)$
                \State $\hat{m} \leftsample A^{\mathcal{O}_\texttt{ATK}}(1^\lambda, \pk, c^\ast)$
                \State \Return $\llbrack \hat{m} = m^\ast \rrbrack$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}\hspace{0.05\textwidth}
    \begin{minipage}[t]{0.4\textwidth}
        \begin{algorithm}[H]
            \caption*{\texttt{IND-ATK} Game}
            \begin{algorithmic}[1]
                \State $(\pk, \sk) \leftsample \keygen(1^\lambda)$
                \State $(m_0, m_1) \leftsample A^{\mathcal{O}_\texttt{ATK}}(1^\lambda, \pk)$
                \State $b \leftsample \{0,1\}$
                \State $c^\ast \leftsample \encrypt(\pk, m_b)$
                \State $\hat{b} \leftsample A^{\mathcal{O}_\texttt{ATK}}(1^\lambda, \pk, c^\ast)$
                \State \Return $\llbrack \hat{b} = b \rrbrack$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}
    \caption{The one-way game, indistinguishability game, plaintext-checking oracle (PCO), and decryption oracle. $\texttt{ATK} \in \{\texttt{CPA}, \texttt{PCA}, \texttt{CCA}\}$}
\end{figure}

Depending on the attack model, the adversary may have access to various oracles. In public-key cryptography, adversaries are always assumed to have the public key with which they can mount chosen-plaintext attack. If the adversary has access to a plaintext-checking oracle $\pco$ then it can mount plaintext-checking attack (PCA). Where the adversary has access to a decryption oracle, it can mount chosen-ciphertext attacks (CCA). A PKE is \texttt{OW-ATK}/\texttt{IND-ATK} secure if no efficient adversary with access to the corresponding oracle(s) can win the corresponding game with non-negligible advantage.

@@

The decapsulation oracle $\mathcal{O}^\decap$ takes a ciphertext $c$ and returns the output of the $\decap$ routine using the secret key. The advantage $\epsilon_\texttt{IND-CCA}$ of an IND-CCA adversary $\mathcal{A}_\texttt{IND-CCA}$ is defined by

@@

\subsection{Message authentication code (MAC)}\label{sec:message-authentication-code}
A message authentication code $\mac$ is a collection of routines $(\sign, \verify)$ defined over some key space $\mathcal{K}$, some message space $\mathcal{M}$, and some tag space $\mathcal{T}$. The signing routine $\sign(k, m)$ takes the secret key $k \in \mathcal{K}$ and some message, and outputs a tag $t$. The verification routine $\verify(k, m, t)$ takes the triplet of secret key, message, and tag, and outputs $\texttt{1}$ if the message-tag pair is valid under the secret key, or $\texttt{0}$ otherwise. Many MAC constructions are deterministic. For these constructions it is simpler to denote the signing routine by $t \leftarrow \mac(k, m)$ and perform verification using a simple comparison.

@@

\begin{figure}[H]
    \centering
    \begin{minipage}{0.7\textwidth}
        \begin{algorithm}[H]
            \caption*{\texttt{EUF-CMA} game}

            \begin{algorithmic}[1]
                \State $k^\ast \leftsample \mathcal{K}$
                \State $(\hat{m}, \hat{t}) \leftsample \mathcal{A}^{\mathcal{O}_\sign}()$
                \State \Return $
                    \llbrack \verify(k^\ast, \hat{m}, \hat{t}) 
                    \land (\hat{m}, \hat{t}) \not\in \mathcal{O}_\sign
                    \rrbrack
                $
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}
    \caption{The existential forgery game}\label{fig:euf-cma-game}
\end{figure}

The advantage $\adv_\texttt{EUF-CMA}$ of the existential forgery adversary is the probability that it wins the EUF-CMA game.

\section{The encrypt-then-MAC transformation}\label{sec:main-results}
Let $\mathcal{B}^\ast$ denote the set of finite bit strings. Let $\pke(\keygen, \encrypt, \decrypt)$ be a public-key encryption scheme defined over message space $\mathcal{M}$ and ciphertext space $\mathcal{C}$. Let $\mac: \mathcal{K}_\mac \times \mathcal{B}^\ast \rightarrow \mathcal{T}$ be a deterministic message authentication code that takes a key $k \in \mathcal{K}_\mac$, some message $m \in \mathcal{B}^\ast$, and outputs a digest $t \in \mathcal{T}$. Let $G: \mathcal{M} \rightarrow \mathcal{K}_\mac$ be a hash function that maps from $\pke$'s plaintext space to $\mac$'s key space. Let $H: \mathcal{B}^\ast \rightarrow \mathcal{K}_\kem$ be a hash function that maps bit strings into the set of possible shared secrets. The encrypt-then-MAC transformation $\etm[\pke, \mac, G, H]$ constructs a key encapsulation mechanism $\kem_\etm(\keygen_\kem, \encap, \decap)$, whose routines are described in figure \ref{fig:etm-routines}.

\begin{figure}[H]
    \centering
    \begin{minipage}[t]{0.5\textwidth}
        \begin{algorithm}[H]
            \caption*{$\kem_\etm.\keygen()$}
            \begin{algorithmic}[1]
                \State $(\pk, \sk^\prime) \leftsample \pke\texttt{.}\keygen()$
                \State $z \leftsample \mathcal{M}$
                \State $\sk \leftarrow \sk^\prime \Vert z$
                \State \Return $(\pk, \sk)$
            \end{algorithmic}
        \end{algorithm}
        \begin{algorithm}[H]
            \caption*{$\kem_\etm.\encap(\pk)$}
            \begin{algorithmic}[1]
                \State $m \leftsample \mathcal{M}$
                \State $k \leftarrow G(m)$
                \State $c^\prime \leftsample \pke\texttt{.}\encrypt(\pk, m)$
                \State $t \leftarrow \mac(k, c^\prime)$
                \State $K \leftarrow H(m, c^\prime)$
                \State $c \leftarrow c^\prime \Vert t$
                \State \Return $(c, K)$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}\hfill
    \begin{minipage}[t]{0.49\textwidth}
        \begin{algorithm}[H]
            \caption*{$\kem_\etm.\decap(\sk, c)$}
            \begin{algorithmic}[1]
                \State $(c^\prime, t) \leftarrow c$
                \State $(\sk^\prime, z) \leftarrow \sk$
                \State $\hat{m} \leftarrow \pke\texttt{.}\decrypt(\sk^\prime, c^\prime)$
                \State $\hat{k} \leftarrow G(\hat{m})$
                \If{$\mac(\hat{k}, c^\prime) \neq t$}
                    \State $K \leftarrow H(z, c^\prime)$
                \Else
                    \State $K \leftarrow H(\hat{m}, c^\prime)$
                \EndIf
                \State \Return $K$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}
    \caption{$\kem_\etm$ routines}\label{fig:etm-routines} 
\end{figure}

@@

\begin{figure}[H]
    \centering
    \begin{minipage}{0.6\textwidth}
    \begin{algorithm}[H]
        \caption*{$\pco(m, c)$}
        \begin{algorithmic}[1]
            \State $k \leftarrow G(m)$
            \State $t \leftarrow \mac(k, c)$
            \State \Return $\llbrack \mathcal{O}^\decap((c, t)) = H(m, c)\rrbrack$
        \end{algorithmic}
    \end{algorithm}
    \end{minipage}
    \caption{Every decapsulation oracle can be converted into a plaintext-checking oracle}\label{fig:pco-using-decap-oracle}
\end{figure}

However, a decapsulation oracle can still give out some information: for a known plaintext $m$, all possible encryptions $c^\prime \leftsample \encrypt(\pk, m)$ can be correctly signed, while ciphertexts that don't decrypt back to $m$ cannot be correctly signed. This means that a decapsulation oracle can be converted into a plaintext-checking oracle (see figure \ref{fig:pco-using-decap-oracle}), so every chosen-ciphertext attack against the KEM can be converted into a plaintext-checking attack against the underlying PKE.

On the other hand, if the underlying PKE is one-way secure against plaintext-checking attack that makes $q$ plaintext-checking queries, then encrypt-then-MAC KEM is semantically secure under chosen ciphertext attacks making the same number of decapsulation queries:

@@

\subsection{Proof of theorem \ref{thm:ow-pca-implies-kem-ind-cca2}}
We will prove theorem \ref{thm:ow-pca-implies-kem-ind-cca2} using a sequence of game. A summary of the the sequence of games can be found in figure \ref{fig:etm-ind-cc2-sequence-of-games} and \ref{fig:ow-pca-simulates-game-3}. From a high level we made three incremental modifications to the IND-CCA game for $\kem_\etm$: replace true decapsulation with simulated decapsulation, replace the pseudorandom MAC key $k^\ast \leftarrow G(m^\ast)$ with a truly random MAC key $k^\ast \leftsample \mathcal{K}_\mac$, and finally replace pseudorandom shared secret $K_0 \leftarrow H(m^\ast, c^\prime)$ with a truly random shared secret $K_0 \leftsample \mathcal{K}_\kem$. A OW-PCA adversary can then simulate the modified IND-CCA game for the KEM adversary, and the advantage of the OW-PCA adversary is associated with the probability of certain behaviors of the KEM adversary.

>>>>

\subsection{Our contribution}
In this paper, we propose a generic IND-CCA secure KEM using techniques adapted from DHIES. We call our proposed scheme the encrypt-then-MAC transformation due to the conceptual similarity to the namesake symmetric encryption technique for achieving authenticated encryption. Our scheme mainly differs from DHIES in its versatility and input requirement. Whereas the IND-CCA security of DHIES reduces specifically to the Gap Diffie-Hellman assumption, the chosen-ciphertext security of the encrypt-then-MAC KEM reduces more generally to the OW-PCA security \cite{DBLP:conf/ctrsa/OkamotoP01} of the input scheme (we will show in Section \ref{sec:main-results} that the Gap Diffie-Hellman assumption implies OW-PCA security). In addition, we propose that because each call to encapsulation samples a fresh PKE plaintext, the encrypt-then-MAC KEM can be instantiated with one-time secure MAC such as Poly1305 (M. Abdalla et al. originally proposed to use HMAC and CBC-MAC, which are many-time secure MAC but less efficient than one-time MAC, see Section \ref{sec:choosing-a-message-authenticator}), which greatly improves the computational efficiency of the scheme.

@@

\textbf{Security.} In Section \ref{sec:main-results}, we prove that if the input PKE is one-way secure against plaintext-checking attack (OW-PCA) and the MAC is one-time existentially unforgeable, then the output KEM is IND-CCA secure. The IND-CCA security of the output KEM reduces tightly to the OW-PCA security of the input PKE, and non-tightly to the unforgeability of the MAC, though the non-tightness will likely not be a weakness because the construction makes it easy to plug in MAC of any desired security level.

\textbf{Implementation and performance comparison.} In Section \ref{sec:practical-instantiation-with-mlkem} and \ref{sec:performance-comparison}, we applied the encrypt-then-MAC transformation to the K-PKE subroutines defined in ML-KEM \cite{FIPS203}, which we call $\mlkemplus$. Compared with ML-KEM, which is based on the $U^{\not\bot}$ variant (implicit rejection with rigidity) of the modular Fujisaki-Okamoto KEM transformation, $\mlkemplus$ demonstrates massive performance improvements in decapsulation while only incurring minimal performance penalty in encapsulation runtime and ciphertext size. On an AMD EPYC 9R14 CPU, $\mlkemplus$ with Poly1305 achieves between 72\%-80\% reduction of CPU cycles needed for decapsulation while only incurring 2\%-7\% increase of CPU cycles needed for encapsulation. See Table \ref{tbl:cpu-cycles-summary} for a summary of performance comparison of individual routines against ML-KEM.

@@

    \caption{$\mlkemplus$ with Poly1305 achieves large performance improvements in decapsulation at the cost of minimal penalty in encapsulation performance and ciphertext size}\label{tbl:cpu-cycles-summary}
\end{table}

We also measured the round trip time of key exchange protocols with various modes of authentication. When compared to ML-KEM, $\mlkemplus$ achieves 24\%-28\% reduction of round trip time in unauthenticated key exchange (KE), 29\%-35\% in unilaterally authenticated key exchange (UAKE), and 35\%-48\% reduction in mutually authenticated key exchange (AKE). See Table \ref{tbl:rtt-summary} for a summary of round trip times.

@@

    \begin{minipage}[t]{0.34\textwidth}
        \begin{algorithm}[H]
            \caption*{$\kem^{\not\bot}_m.\keygen()$}
            \begin{algorithmic}[1]
                \State $(\pk, \sk^\prime) \leftsample \texttt{PKE.KeyGen}()$
                \State $z \leftsample \mathcal{M}$
                \State $\sk \leftarrow (\sk^\prime, \pk, z)$
                \State \Return $(\pk, \sk)$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}
    \begin{minipage}[t]{0.3\textwidth}
        \begin{algorithm}[H]
            \caption*{$\kem^{\not\bot}_m.\encap(\pk)$}
            \begin{algorithmic}[1]
                \State $m \leftsample \mathcal{M}$
                \State $r \leftarrow G(m)$
                \State $c \leftarrow \texttt{PKE.Enc}(\pk, m ,r)$
                \State $K \leftarrow H(m)$
                \State \Return $(c, K)$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}
    \begin{minipage}[t]{0.3\textwidth}
        \begin{algorithm}[H]
            \caption*{$\kem^{\not\bot}_m.\decap(\sk, c)$}
            \begin{algorithmic}[1]
                \State $\hat{m} \leftarrow \texttt{PKE.Dec}(\sk^\prime, c)$
                \State $\hat{r} \leftarrow G(m)$
                \State $\hat{c} \leftarrow \texttt{PKE.Enc}(\pk, \hat{m}, \hat{r})$
                \If{$\hat{c} = c$}
                    \State $K \leftarrow H(\hat{m})$
                \Else
                    \State $K \leftarrow H(z, c)$
                \EndIf
                \State \Return $K$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}

    \caption{The $U^{\not\bot}_m$ variant of Fujisaki-Okamoto transformation is used in ML-KEM}\label{fig:u-notbot-m-routines}
\end{figure}

The modular Fujisaki-Okamoto KEM transformation is remarkably successful because of the simplicity of its construction, the tightness of the security bound, and the proven (though non-tight) security against quantum adversaries. It was adopted by many submissions to NIST's post-quantum cryptography competition, including Kyber \cite{DBLP:conf/eurosp/BosDKLLSSSS18}, Saber \cite{DBLP:conf/africacrypt/DAnversKRV18}, FrodoKEM \cite{DBLP:conf/ccs/BosCDMNNRS16}, and classic McEliece \cite{classicmceliecespec} among others. The $U^{\not\bot}_m$ variant of the Fujisaki-Okamoto transformation (see Figure \ref{fig:u-notbot-m-routines}) is adopted in ML-KEM.

However, the Fujisaki-Okamoto transformation is not perfect. Because it uses de-randomization and re-encryption for achieving ciphertext non-malleability, the Fujisaki-Okamoto transformation suffers from the following two problems: \begin{itemize}
    \item \textbf{Computational inefficiency}. The decapsulation routine needs to re-encrypt the decryption to ensure ciphertext has not been tempered with. For input PKE whose encryption routine carries substantial computational cost, such as most lattice-based cryptosystems, re-encryption slows down decapsulation significantly.
    \item \textbf{Side-channel vulnerability}. Re-encryption also introduces side-channels that can leak information about the decrypted PKE plaintext. As demonstrated in \cite{ueno2022curse}\cite{ravi2019generic}\cite{DBLP:journals/tches/TanakaUXITH23}, these side-channels can be converted into efficient plaintext-checking attacks that can fully recover the secret key.
\end{itemize}

@@

\subsection{Public-key encryption scheme}
\textbf{Syntax.} A public-key encryption scheme $\pke(\keygen, \encrypt, \decrypt)$ is a collection of three routines defined over some plaintext space $\mathcal{M}$ and some ciphertext space $\mathcal{C}$. Key generation $(\pk, \sk) \leftsample \keygen()$ is a randomized routine that returns a keypair. The encryption routine $\encrypt: (\pk, m) \mapsto c$ encrypts the input plaintext under the input public key. The decryption routine $\decrypt: (\sk, c) \mapsto m$ decrypts the input ciphertext under the input secret key. Where the encryption routine is randomized, we denote the randomness by a coin $r \in \mathcal{R}$, where $\mathcal{R}$ is called the coin space. The decryption routine is assumed to always be deterministic. Some decryption routines can detect malformed ciphertext and output the rejection symbol $\bot$ accordingly.

@@

Where the expectation is taken with respect to the probability distribution of all possible keypairs $(\pk, \sk) \leftsample \texttt{PKE.KeyGen()}$. For many lattice-based cryptosystems, including ML-KEM, decryption failures could leak information about the secret key, although the probability of a decryption failure is low enough that classical adversaries cannot exploit decryption failure more than they can defeat the underlying lattice problem.

\textbf{Security.} The security of public-key encryption is conventionally discussed within the context of adversarial games played between a challenger and an adversary \cite{DBLP:conf/stoc/GoldwasserM82}. There are two main types of games: in the one-wayness game, the adversary is given a random encryption, then asked to produce the correct decryption; in the indistinguishability game, the adversary is given the encryption of one of two adversary-chosen plaintexts, then asked to decide which of the plaintexts corresopnds with the given encryption. Depending on the attack model, the adversary may have access to various oracles. Within the context of public-key cryptography, adversaries are always assumed to have the public key with which they can mount chosen-plaintext attack (CPA). If the adversary has access to a plaintext-checking oracle (PCO) \cite{DBLP:conf/ctrsa/OkamotoP01} then it can mount plaintext-checking attack (PCA). Where the adversary has access to a decryption oracle, it can mount chosen-ciphertext attacks (CCA).

@@

    \begin{minipage}[t]{0.3\textwidth}
        \begin{algorithm}[H]
            \caption*{\texttt{OW-ATK} Game}
            \begin{algorithmic}[1]
                \State $(\pk, \sk) \leftsample \keygen(1^\lambda)$
                \State $m^\ast \leftsample \mathcal{M}$
                \State $c^\ast \leftsample \encrypt(\pk, m^\ast)$
                \State $\hat{m} \leftsample A^{\mathcal{O}_\texttt{ATK}}(1^\lambda, \pk, c^\ast)$
                \State \Return $\llbrack \hat{m} = m^\ast \rrbrack$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}
    \begin{minipage}[t]{0.33\textwidth}
        \begin{algorithm}[H]
            \caption*{\texttt{IND-ATK} Game}
            \begin{algorithmic}[1]
                \State $(\pk, \sk) \leftsample \keygen(1^\lambda)$
                \State $(m_0, m_1) \leftsample A^{\mathcal{O}_\texttt{ATK}}(1^\lambda, \pk)$
                \State $b \leftsample \{0,1\}$
                \State $c^\ast \leftsample \encrypt(\pk, m_b)$
                \State $\hat{b} \leftsample A^{\mathcal{O}_\texttt{ATK}}(1^\lambda, \pk, c^\ast)$
                \State \Return $\llbrack \hat{b} = b \rrbrack$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}
    \begin{minipage}[t]{0.33\textwidth}
        \begin{algorithm}[H]
            \caption*{$\mathcal{O}_\pco(m, c)$}
            \begin{algorithmic}[1]
                \State \Return $\llbrack m = \decrypt(\sk, c) \rrbrack$
            \end{algorithmic}
        \end{algorithm}
        \begin{algorithm}[H]
            \caption*{$\mathcal{O}_\decrypt(c)$}
            \begin{algorithmic}[1]
                \State \Return $\decrypt(\sk, c)$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}
    \caption{The one-way game, indistinguishability game, plaintext-checking oracle (PCO), and decryption oracle. $\texttt{ATK} \in \{\texttt{CPA}, \texttt{PCA}, \texttt{CCA}\}$}
\end{figure}

The advantage of an adversary in the \texttt{OW-ATK} game is the probability that it ouputs the correct decryption. The advantage of an adversary in the \texttt{IND-ATK} game is defined below. A PKE is \texttt{OW-ATK}/\texttt{IND-ATK} secure if no efficient adversary has non-negligible advantage in the corresponding security game.

\begin{equation*}
    \texttt{Adv}_\texttt{IND-ATK}(A) = \left\vert P\left[ \hat{b} = b \right] - \frac{1}{2} \right\vert
\end{equation*}

@@

The decapsulation oracle $\mathcal{O}^\decap$ takes a ciphertext $c$ and returns the output of the $\decap$ routine using the secret key. The advantage $\epsilon_\texttt{IND-CCA}$ of an IND-CCA adversary $\mathcal{A}_\texttt{IND-CCA}$ is defined by the adversary's ability to correctly distinguish the two cases beyond blind guess:

@@

A KEM is \texttt{IND-ATK} secure if no efficient adversary has non-negligible advantage in the corresponding security game.

\subsection{Message authentication code (MAC)}\label{sec:message-authentication-code}
A message authentication code $\mac(\keygen, \sign, \verify)$ is a collection of routines defined over some key space $\mathcal{K}$, some message space $\mathcal{M}$, and some tag space $\mathcal{T}$. The signing routine $\sign(k, m)$ authenticates the message $m$ under the secret key $k$ by producing a tag $t$ (also called digest). The verification routine $\verify(k, m, t)$ takes the triplet of secret key, message, and tag, and outputs $\texttt{1}$ if the message-tag pair is valid under the secret key, or $\texttt{0}$ otherwise. Many MAC constructions are deterministic. For these constructions it is simpler to denote the signing routine by $t \leftarrow \mac(k, m)$ and perform verification using a simple comparison.

@@

\begin{figure}[H]
    \centering
    \begin{minipage}{0.53\textwidth}
        \begin{algorithm}[H]
            \caption*{\texttt{EUF-CMA} game}
            \begin{algorithmic}[1]
                \State $k^\ast \leftsample \mathcal{K}$
                \State $(\hat{m}, \hat{t}) \leftsample \mathcal{A}^{\mathcal{O}_\sign}()$
                \State \Return $
                    \llbrack \verify(k^\ast, \hat{m}, \hat{t}) 
                    \land (\hat{m}, \hat{t}) \not\in \mathcal{O}_\sign
                    \rrbrack
                $
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}
    \caption{The existential forgery game}\label{fig:euf-cma-game}
\end{figure}

The advantage $\adv_\texttt{EUF-CMA}$ of the existential forgery adversary is the probability that it wins the EUF-CMA game. Some MACs are one-time existentially unforgeable, meaning that each secret key can be used to authenticate only a single message. The corresponding security game is modified such that the signing oracle will only answer a single signing query.

\section{The encrypt-then-MAC transformation}\label{sec:main-results}
In this section we introduce the encrypt-then-MAC transformation that transforms an OW-PCA secure PKE and an one-time existentially unforgeable MAC into an IND-CCA secure KEM. In Section \ref{sec:proof-of-theorem} we reduce the IND-CCA security of the KEM tightly to the OW-PCA security of the underlying PKE, and non-tightly to the unforgeability of the MAC. In Section \ref{sec:elgamal-is-ow-pca}, we show that the ``Hashed ElGamal'' cryptosystem in \cite{DBLP:conf/ctrsa/AbdallaBR01} is a special case of the encrypt-then-MAC transformation by reducing the OW-PCA security of the ElGamal cryptosystem to the Gap Diffie-Hellman assumption.

Let $\mathcal{B}^\ast$ denote the set of finite bit strings. Let $\pke(\keygen, \encrypt, \decrypt)$ be a public-key encryption scheme defined over message space $\mathcal{M}$ and ciphertext space $\mathcal{C}$. Let $\mac: \mathcal{K}_\mac \times \mathcal{B}^\ast \rightarrow \mathcal{T}$ be a deterministic message authentication code that takes a key $k \in \mathcal{K}_\mac$, some message $m \in \mathcal{B}^\ast$, and outputs a digest $t \in \mathcal{T}$. Let $G: \mathcal{M} \rightarrow \mathcal{K}_\mac$ be a hash function that maps from $\pke$'s plaintext space to $\mac$'s key space. Let $H: \mathcal{B}^\ast \rightarrow \mathcal{K}_\kem$ be a hash function that maps bit strings into the set of possible shared secrets. The encrypt-then-MAC transformation $\etm[\pke, \mac, G, H]$ constructs a key encapsulation mechanism $\kem_\etm(\keygen, \encap, \decap)$, whose routines are described in Figure \ref{fig:etm-routines}.

\begin{figure}[h]
    \centering
    \begin{minipage}[t]{0.5\textwidth}
        \begin{algorithm}[H]
            \caption*{$\kem_\etm.\keygen()$}
            \begin{algorithmic}[1]
                \State $(\pk, \sk^\prime) \leftsample \pke\texttt{.}\keygen()$
                \State $z \leftsample \mathcal{M}$
                \State $\sk \leftarrow (\sk^\prime, z)$
                \State \Return $(\pk, \sk)$
            \end{algorithmic}
        \end{algorithm}
        \begin{algorithm}[H]
            \caption*{$\kem_\etm.\encap(\pk)$}
            \begin{algorithmic}[1]
                \State $m \leftsample \mathcal{M}$
                \State $k \leftarrow G(m)$
                \State $c^\prime \leftsample \pke\texttt{.}\encrypt(\pk, m)$
                \State $t \leftarrow \mac(k, c^\prime)$
                \State $K \leftarrow H(m, c^\prime)$
                \State $c \leftarrow (c^\prime, t)$
                \State \Return $(c, K)$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}\hfill
    \begin{minipage}[t]{0.49\textwidth}
        \begin{algorithm}[H]
            \caption*{$\kem_\etm.\decap(\sk, c)$}
            \begin{algorithmic}[1]
                \State $(c^\prime, t) \leftarrow c$
                \State $(\sk^\prime, z) \leftarrow \sk$
                \State $\hat{m} \leftarrow \pke\texttt{.}\decrypt(\sk^\prime, c^\prime)$
                \State $\hat{k} \leftarrow G(\hat{m})$
                \If{$\mac(\hat{k}, c^\prime) \neq t$}
                    \State $K \leftarrow H(z, c^\prime)$
                \Else
                    \State $K \leftarrow H(\hat{m}, c^\prime)$
                \EndIf
                \State \Return $K$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}
    \caption{The encrypt-then-MAC transformation builds a KEM $\kem_\etm$ using a $\pke(\keygen, \encrypt, \decrypt)$, a $\mac$, and two hash functions $G, H$}\label{fig:etm-routines} 
\end{figure}

@@

However, a decapsulation oracle can still give out some information: for a known plaintext $m$, all possible encryptions $c^\prime \leftsample \encrypt(\pk, m)$ can be correctly signed, while ciphertexts that don't decrypt back to $m$ cannot be correctly signed. This means that a decapsulation oracle can be converted into a plaintext-checking oracle (see Figure \ref{fig:pco-using-decap-oracle}), so every chosen-ciphertext attack against the KEM can be converted into a plaintext-checking attack against the underlying PKE.

\begin{figure}[H]
    \centering
    \begin{minipage}{0.6\textwidth}
    \begin{algorithm}[H]
        \caption*{$\pco(m, c)$}
        \begin{algorithmic}[1]
            \State $k \leftarrow G(m)$
            \State $t \leftarrow \mac(k, c)$
            \State \Return $\llbrack \mathcal{O}^\decap((c, t)) = H(m, c)\rrbrack$
        \end{algorithmic}
    \end{algorithm}
    \end{minipage}
    \caption{With the encrypt-then-MAC transformation, every chosen-ciphertext attack against the KEM can be converted into a plaintext-checking attack against the underlying PKE}\label{fig:pco-using-decap-oracle}
\end{figure}

On the other hand, if the underlying PKE is OW-PCA secure and the underlying MAC is one-time existentially unforgeable, then the encrypt-then-MAC KEM is IND-CCA secure:

@@

\subsection{Proof of theorem \ref{thm:ow-pca-implies-kem-ind-cca2}}\label{sec:proof-of-theorem}
We will prove theorem \ref{thm:ow-pca-implies-kem-ind-cca2} using a sequence of game. A summary of the the sequence of games can be found in Figure \ref{fig:etm-ind-cc2-sequence-of-games} and \ref{fig:ow-pca-simulates-game-3}. From a high level we made three incremental modifications to the IND-CCA game for $\kem_\etm$: replace true decapsulation with simulated decapsulation, replace the pseudorandom MAC key $k^\ast \leftarrow G(m^\ast)$ with a truly random MAC key $k^\ast \leftsample \mathcal{K}_\mac$, and finally replace pseudorandom shared secret $K_0 \leftarrow H(m^\ast, c^\prime)$ with a truly random shared secret $K_0 \leftsample \mathcal{K}_\kem$. A OW-PCA adversary can then simulate the modified IND-CCA game for the KEM adversary, and the advantage of the OW-PCA adversary is associated with the probability of certain behaviors of the KEM adversary.
